Ore Polynomials in Sage

### 2013-06-10 paper/main.tex
paper: minor modifications.

@@

@@

@@

\subsection{to list?}

-- available for shift and q-shift algebras.

@@

\begin{sageexample}
  sage: R.<n> = ZZ[]; RS.<Sn> = OreAlgebra(R, 'Sn')
  sage: inverse_factorials = (n + 1) * Sn - 1
  sage: partial_sums = inverse_factorials.annihilator_of_sum()
  sage: partial_sums
\end{sageexample}

@@

In some cases (for example when the base ring is $\mathbb{Z}$ or
$\mathbb{Z}[x]$), isolated values can be
computed asymptotically faster for large $n$ using the binary splitting
technique.
The \verb|forward_matrix_bsplit| method, called with argument $n$,
returns a matrix $P$ and a denominator $Q$ such that $P / Q$ multiplied a
column vector of initial values $c_0, c_1, \ldots$
yields $c_n, c_{n+1}, \ldots$.
This way, computing $10^5$ digits of $e$ takes a fraction of a second:

@@

@@

Guessing is a technique for detecting patterns in sequences of numbers, which is
often used in combinatorics. It is one of the most popular features of packages
like gfun, and there are even some special purpose packages dedicated to this
technique~\cite{kauers09a,hebisch11}.  The basic idea is simple. If we are given
a finite array of numbers, though of as the first terms of an infinite sequence,
we may want to know whether this sequence satisfies a recurrence. The algorithm
behind a guessing engine searches for small equations matching the given data.
Generically, no such equations exist, so if some are found, it is fair to
``guess'' that they are in fact valid equations for the whole infinite sequence.

>>>>

/// disclaimer: we don't explain algorithms here, but only the software ///

@@

-- conceptually the base ring is always a field, even when it's not.

@@

-- merge with previous section?

\subsection{conversions?}

-- re2de and all its relatives

@@

\subsection{Computation of Sequence Terms}

(... a few introductory words ... available for shift and q-shift algebras ...)

@@

\begin{sageexample}
  sage: R.<n> = ZZ[]; RS.<Sn> = OreAlgebra(R)
  sage: inverse_factorials = (n + 1) * Sn - 1
  sage: partial_sums = inverse_factorials.annihilator_of_sum()
  sage: partial_sums
\end{sageexample}

@@

In some cases (for example when the base ring is $\mathbb{Z}$ or
$\mathbb{Z}[x]$), isolated values can be
computed asymptotically faster for large $n$ using the binary splitting
technique.
The \verb|forward_matrix_bsplit| method, called with argument~$n$,
returns a matrix~$P$ and a polynomial~$Q$ such that $P / Q$ multiplied by a
column vector of initial values $c_0, c_1, \ldots$
yields $c_n, c_{n+1}, \ldots$.
This way, computing $10^5$ digits of $\mathrm{e}$ takes a fraction of a second:

@@

... add a figure with timings? ...

@@

Guessing is, in some sense, the reverse operation of \verb|to_list|. It is one
of the most popular features of packages like gfun, and there are even some
special purpose packages dedicated to this technique~\cite{kauers09a,hebisch11}.
The basic idea is simple. If we are given a finite array of numbers, thought of
as the first terms of an infinite sequence, we may want to know whether this
sequence satisfies a recurrence. The algorithm behind a guessing engine searches
for small equations matching the given data.  Generically, no such equations
exist, so if some are found, it is fair to ``guess'' that they are in fact valid
equations for the whole infinite sequence.
